---
title: "Significant Volcanoes"
author: "Martin Hill"
date: "01 April 2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
```

## Souce Material

Data provided care of the National Geophysical Data Center, NOAA.

https://www.ngdc.noaa.gov/nndc/servlet/ShowDatasets?dataset=102557&search_look=50&display_look=50

Their database globally lists over 500 significant eruptions which includes information on the latitude, longitude, elevation, type of volcano, and last known eruption.

## Criteria
A significant eruption meets one of the folowing crieria:  

- Caused fatalities
- Caused moderate damage (approximately $1 million or more)
- With a Volcanic Explosivity Index (VEI) of 6 or larger
- Caused a tsunami
- Was associated with a major earthquake

For a complete list of current and past activity for all volcanoes on the planet active during the last 10,000 years, please see Smithsonian Institution's Global Volcanism Program (GVP). 


```{r loaddata, cache=TRUE}
download.file(destfile = "volerup.txt",
      url = "https://www.ngdc.noaa.gov/nndc/struts/results?type_0=Exact&query_0=$HAZ_EVENT_ID&t=102557&s=50&d=54&dfn=volerup.txt")

df<-read.delim2(file="volerup.txt", stringsAsFactors = FALSE)
#df<-df[df$Country == "Japan",]
#set col to VEI, where NA to be -1 and make as index
df$VEIsize<-df$VEI
df["VEIsize"][is.na(df["VEIsize"])] <- -1
df$VEIsize<-df$VEIsize+2
class(df$Latitude)<-"numeric"
class(df$Longitude)<-"numeric"
```

## Significant Eruptions
Here the size indicates the VEI. The Agents are assigned letters (A,E,F,G,I,L,M,P,S,T,W).
```{r Map, echo=FALSE, eval=TRUE }
library(leaflet)
YearCol <- function(Year) {
      if(Year < -4000) "grey"
      else if(Year < -3000) "violet"
      else if(Year < -2000) "indigo"
      else if(Year < -1000) "blue"
      else if(Year < 0) "green" #1st millenium BC
      else if(Year < 1000) "yellow" #1st millenium AD
      else if(Year < 2000) "orange"
      else if(Year < 3000) "red"
      else "black"
}

df$Yearcolour <- sapply(df$Year, function(x) YearCol(x))

Yearcolour <- sapply(c(-5000,-4000,-3000,-2000,-1000,0,1000,2000), function(x) YearCol(x))
YearLabel <-c("5000 BC- 4001 BC","4000 BC- 3001 BC","3000 BC- 2001 BC","2000 BC- 1001 BC","1000 BC- 1 BC", "1 AD- 1000 AD", "1001 AD- 2000 AD","2001 AD- 3000 AD")

map<-data.frame(lat = df$Latitude,
                lng = df$Longitude,
                col = sapply(df$Year, function(x) YearCol(x)),
                size = df$VEIsize*2,
                stringsAsFactors = TRUE
                )
content <- paste(sep = "<br/>",
  df$Name,
  paste(sep = "", "Year : ", as.character(df$Year)),
  paste(sep = "", "Elevation : ", as.character(df$Elevation)),
  paste(sep = "", "Agent : ", df$Agent)
)

map%>%leaflet()%>%
      addTiles()%>%
      addCircleMarkers(color = map$col, radius = map$size, popup = content)%>%
      addLegend(labels = YearLabel, colors = Yearcolour)
```


```{r plotly, echo=FALSE, eval=FALSE}
library(plotly)
plot_ly(df, x = ~Longitude, y = ~Latitude, mode = "markers", color = ~Elevation, size = ~VEIsize)
```

##Interactive Map
This link provides a map which can be filtered by a Country, a range of Years and some Agents (or All).
Simply make your selection and press Submit. 

[Interactive Map](https://martinhill001.shinyapps.io/Significant_Volcanoes/)

Assigning: E = Earthquake, P = Pyroclastic Flow, T = Tsunami  


### Example 1, Italy, [Vesuvius](http://www.protezionecivile.gov.it/jcms/en/eruzione_vesuvio_1631.wp)
Year: 1631, Agents P,E,M

### Example 2, United States, [St. Helens](http://volcano.oregonstate.edu/mt-st-helens)
Year: 1980, Agents P,M,T


